---
title: "day10-activities"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(janitor)
library(here)
library(sf)
```

## **Your task:**

-   Download both datasets

-   Explore the metadata for each to ensure you know what they contain, variables, units, etc. **NOTE:** How are missing values stored in the ice duration variable?

-   Create a new version-controlled R Project

-   In a single R Markdown (or Quarto) document:

    -   Read in both files (ice cover duration and meteorological data). Note that the ice cover data has some NA values stored as -999.

    ```{r}
    ice_cover_duration <- read_csv(here("data","ntl33_v7.csv"), na = "-999")

    meteorological_data <- read_csv(here("data", "ntl20_v6.csv"), na = "-999")
    ```

    -   Create a visualization that compares ice duration across different lakes in the Madison Lake Area. Include both the actual ice duration values *and* some representation of summary statistics for each lake. Careful if you use a boxplot over a jitter or beeswarm - what values might be doubly represented? Can you find an argument that would help you remove those from the boxplot?

    ```{r}
    ice_subset1 <- ice_cover_duration %>% 
      group_by(lakeid) %>% 
      summarise(ice_mean = mean(ice_duration, na.rm = TRUE))
    ggplot(ice_cover_duration, aes(y = ice_duration , x = lakeid))+
      geom_jitter()
    ```

    -   Find the mean ice cover duration in Madison Lake Area by year (i.e., you\'ll only be grouping by the year variable here). Understanding check: what are you finding the mean of here? Why are there multiple observations for each year?)

    ```{r}
    ice_subset2 <- ice_cover_duration %>% 
      group_by(year4) %>% 
      summarise(ice_mean = mean(ice_duration, na.rm = TRUE))
    ```

    -   Create an exploratory visualization of mean ice cover duration by year for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

    ```{r}
    ice_cover_duration %>% 
      group_by(year4) %>% 
      summarise(ice_mean = mean(ice_duration, na.rm = TRUE))
    ggplot(ice_cover_duration, aes(y = ice_duration , x = year4, color = lakeid))+
      geom_jitter()
    ```

    -   Find the mean air temperature (using the daily average air temperature - adjusted) in Madison Lake Area **by year** using **only observations from winter months (December, January, February)**

    ```{r}

    ```

    -   Create an exploratory visualization of mean winter temperatures for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

    ```{r}

    ```

    -   Join the mean winter air temperatures to the mean ice cover duration data you found above

    ```{r}

    ```

    -   Create an exploratory scatterplot of mean winter temperatures versus mean ice duration. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

    ```{r}

    ```

    -   Explore thaw dates since 1970. Create a subset of the ice cover duration data since 1970 (ok to include 1970), then convert the `ice_off` column to a `Date`. Use `lubridate::yday()` to pull the numeric *day of the year* from that Date. Create an exploratory plot of the ice off day (numeric day-of-the-year) from 1970 - present. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.
